Simulation study in rectangular domain

### 2024-04-11 R/simulation_study/simulation_study_rect.Rmd
update simulation study

---
title: "Simulation study in rectangular domain"
output: html_notebook
---

@@

#some auxiliary functions
source(file.path("/home","delporta","Documents","Recherche","Codes","CVM","CVM_functions.R"))
```

# Create the rectangle

@@

#rectangular domain 
large_rectangle_coords <- matrix(c(0, 0, 0, 10000, 10000, 10000, 10000, 0, 0, 0), ncol = 2, byrow = TRUE)
small_rectangle_coords <- matrix(c(1000,1000, 1000, 9000, 9000, 9000, 9000, 1000, 1000, 1000), ncol = 2, byrow = TRUE)

@@

# Sample initial positions uniformly in the domain

@@

#number of tracks
n_samples=100

@@

# Fix time steps for observations

@@

hf=1/60 #1 minute time step
lf=5/60 #5 minute time step

@@

#low frequency time step
times_lf=seq(0,Tmax,by=lf)
n_lf=length(times_lf)-1
```

# Set observation error

```{r}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.003
H0_lf=array(rep(sigma_obs_0^2*diag(2),n_lf),dim=c(2,2,n_lf))
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf),dim=c(2,2,n_hf))
#GPS measurement error 
sigma_obs_1=0.03
H1_lf=array(rep(sigma_obs_1^2*diag(2),n_lf),dim=c(2,2,n_lf))
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf),dim=c(2,2,n_hf))

@@

# Smooth parameters of the CRCVM

Define smooth functions for tau and omega.

@@

ftau=function(Dshore,theta,D0=0.3,delta0=0.2,tau0=1,tau1=2,tau2=2,lambda=4,kappa=0.1) {
  coeff=exp(-kappa*(Dshore/D0)^2)
  tau=tau0*(1-coeff)+(delta0+(tau0-delta0)*exp(-theta^2/pi^2)+(tau1-delta0)*exp(-lambda*(theta-pi/2)^2)+(tau2-delta0)*exp(-lambda*(theta+pi/2)^2))*coeff
  return (tau)
}

#mixture of independent gaussian
ftau_gauss=function(Dshore,theta,D0=0.5,tau0=1,tau1=1.5) {
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/(pi/10))^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/(pi/10))^2-(Dshore/D0)^2)
  return (tau)
}

#constant tau
ftau_constant=function(Dshore,theta,tau=1) {
  return (tau)
}

fomega=function(Dshore,theta,D0=0.3,omega0=60*pi,lambda=1.5,kappa=0.5) {
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

Plot smooth functions.
```{r}

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- expand.grid(Dshore,theta)
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_gauss(grid$DistanceShore,grid$theta)
true_par$omega=fomega(grid$DistanceShore,grid$theta)

@@

# Simulation and estimation when only omega is smooth (other parameters are fixed)

## Generate samples : no measurement errors and high frequency

@@

## percentage of samples that reached land

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_ne[data_rect_hf_ne$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count,"percent of the samples reached land")
```

## plot one sample

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_path(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_point(data = data_rect_hf_ne[data_rect_hf_ne$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1* 1000, y = y2 * 1000), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

## Subsample : no measurement errors and low frequency

data_rect_lf_ne=data_rect_hf_ne[seq(1,n_hf,by=lf/hf),]

## plot subsamples

plot_samples_rect_lf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_lf_ne[data_rect_lf_ne$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_path(data=data_rect_lf_ne[data_rect_lf_ne$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_point(data = data_rect_lf_ne[data_rect_lf_ne$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1* 1000, y = y2 * 1000), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggsave(filename="samples_rect_lf_ne_omega.png",plot=plot_samples_rect_lf_ne)
ggplotly(plot_samples_rect_lf_ne)
```

## Generate samples : 30m measurement errors and high frequency

@@

## percentage of samples that reached land

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_e[data_rect_hf_e$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count,"percent of the samples reached land")
```
## plot samples

@@

plot_samples_rect_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_path(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_point(data = data_rect_hf_e[data_rect_hf_e$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1* 1000, y = y2 * 1000), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

@@

## Subsample : measurement errors and low frequency

data_rect_lf_e=data_rect_hf_e[seq(1,n_hf,by=lf/hf),]

## plot subsamples

plot_samples_rect_lf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_lf_e[data_rect_lf_e$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_path(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_point(data = data_rect_hf_e[data_rect_hf_e$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1* 1000, y = y2 * 1000), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggsave(filename="samples_rect_lf_e_omega.png",plot=plot_samples_rect_lf_e)
ggplotly(plot_samples_rect_lf_e)
```
## Compute distance to shore and empirical deviation angles

@@

data_rect_hf_ne=add_covs(data_rect_hf_ne)
data_rect_lf_ne=add_covs(data_rect_lf_ne)

data_rect_hf_e=add_covs(data_rect_hf_e)
data_rect_lf_e=add_covs(data_rect_lf_e)
```

## Estimate from high frequency data without error : ExpShore and tensor splines te

### Fit sde
```{r}

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=c(10,5),bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne1 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1 & data_rect_hf_ne$time<2*24,],type = "RACVM1",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne1$fit(method="BFGS")
```

### Plot smooth parameter omega in 3D

```{r}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
res=crcvm_rect_hf_ne1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne1",npost=1000,level=0.95,xmin=xmin,xmax=xmax,xlabel=xlabel,link=link)
res$plot_omega_theta_ExpShore
```

### Plot true and estimated omega as a function of theta for fixed values of DistanceShore, with confidence intervals.

@@

#plot spline estimates of the parameter omega close and far to the shore

predict_close=data.frame(theta=seq(-pi,pi,length.out=100),ExpShore=rep(1/0.5,100))
predict_far=data.frame(theta=seq(-pi,pi,length.out=100),ExpShore=rep(1/3,100))

sde_par_close <- crcvm_rect_hf_ne1$par(t = "all",new_data=predict_close)
sde_CI_close <- crcvm_rect_hf_ne1$CI_simultaneous(t = "all",new_data=predict_close)
sde_par_far <- crcvm_rect_hf_ne1$par(t = "all",new_data=predict_far)
sde_CI_far <- crcvm_rect_hf_ne1$CI_simultaneous(t = "all",new_data=predict_far)

# Data frame for point estimates
sde_par_df_close <- data.frame(theta=predict_close$theta,omega_estimates = sde_par_close[, "omega"])
sde_par_df_far<- data.frame(theta=predict_far$theta,omega_estimates = sde_par_far[, "omega"])

# Data frame for CIs
sde_ci_df_close<- data.frame(theta=predict_close$theta,
                          lowomega = sde_CI_close["omega", "low",],
                          upomega = sde_CI_close["omega", "upp",])
sde_ci_df_far<- data.frame(theta=predict_far$theta,
                          lowomega = sde_CI_far["omega", "low",],
                          upomega = sde_CI_far["omega", "upp",])

omega_df=data.frame(theta=predict_close$theta,omega_close=fomega(rep(0.5,100),predict_close$theta),omega_far=fomega(rep(3,100),predict_close$theta))

plot_omega_close=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_close) + 
  geom_ribbon(data=sde_ci_df_close,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  geom_line(data=omega_df,aes(theta,omega_close),col="red")+xlab("Theta")+
  ylab("omega")+ggtitle("Distance to shore = 0.5 km")

plot_omega_far=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_far) + 
  geom_ribbon(data=sde_ci_df_far,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  geom_line(data=omega_df,aes(theta,omega_far),col="red")+xlab("Theta")+
  ylab("omega")+ggtitle("Distance to shore = 3 km")

plot_omega_close
plot_omega_far

ggsave(filename="crcvm_rect_hf_ne1/crcvm_rect_hf_ne1_omega_close.png",plot=plot_omega_close)
ggsave(filename="crcvm_rect_hf_ne1/crcvm_rect_hf_ne1_omega_far.png",plot=plot_omega_far)

### Plot true and estimated omega as a function of DistanceShore for fixed values of theta, with confidence intervals.

#plot spline estimates of the parameter omega close and far to the shore

predict_away=data.frame(ExpShore=seq(1/2,1/0.3,length.out=100),theta=rep(pi/3,100))
predict_toward=data.frame(ExpShore=seq(1/2,1/0.3,length.out=100),theta=rep(2*pi/3,100))

sde_par_away <- crcvm_rect_hf_ne1$par(t = "all",new_data=predict_away)
sde_CI_away <- crcvm_rect_hf_ne1$CI_simultaneous(t = "all",new_data=predict_away)
sde_par_toward<- crcvm_rect_hf_ne1$par(t = "all",new_data=predict_toward)
sde_CI_toward <- crcvm_rect_hf_ne1$CI_simultaneous(t = "all",new_data=predict_toward)

# Data frame for point estimates
sde_par_df_away<- data.frame(ExpShore=predict_away$ExpShore,omega_estimates = sde_par_away[, "omega"])
sde_par_df_toward<- data.frame(ExpShore=predict_toward$ExpShore,omega_estimates = sde_par_toward[, "omega"])

# Data frame for CIs
sde_ci_df_away<- data.frame(ExpShore=predict_away$ExpShore,
                          lowomega = sde_CI_away["omega", "low",],
                          upomega = sde_CI_away["omega", "upp",])
sde_ci_df_toward<- data.frame(ExpShore=predict_toward$ExpShore,
                          lowomega = sde_CI_toward["omega", "low",],
                          upomega = sde_CI_toward["omega", "upp",])

omega_df=data.frame(DistanceShore=1/predict_away$ExpShore,omega_away=fomega(1/predict_away$ExpShore,rep(pi/3,100)),
                    omega_toward=fomega(1/predict_toward$ExpShore,rep(2*pi/3,100)))

plot_omega_away=ggplot()+geom_line(aes(1/ExpShore,omega_estimates),data=sde_par_df_away) + 
  geom_ribbon(data=sde_ci_df_away,aes(x=1/ExpShore,ymin=lowomega,ymax=upomega),alpha=0.2)+
  geom_line(data=omega_df,aes(DistanceShore,omega_away),col="red")+xlab("Distance to shore (km)")+
  ylab(TeX(r'(\omega)'))+ggtitle(TeX(r'(\Theta = \pi/3)'))

plot_omega_toward=ggplot()+geom_line(aes(1/ExpShore, omega_estimates),data=sde_par_df_toward) + 
  geom_ribbon(data=sde_ci_df_toward,aes(x=1/ExpShore,ymin=lowomega,ymax=upomega),alpha=0.2)+
  geom_line(data=omega_df,aes(DistanceShore,omega_toward),col="red")+xlab("Distance to shore (km)")+
  ylab("Omega")+ggtitle(TeX(r'(\Theta = 2\pi/3)'))

plot_omega_away
plot_omega_toward

ggsave(filename="crcvm_rect_hf_ne1/crcvm_rect_hf_ne1_omega_away.png",plot=plot_omega_away)
ggsave(filename="crcvm_rect_hf_ne1/crcvm_rect_hf_ne1_omega_toward.png",plot=plot_omega_toward)

### Simulate from fitted model

@@

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_e1$simulate(z0=x0[1,],data=crcvm_rect_hf_e1$data(),atw=atw,land=domain)

@@

@@

plot_sim=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=sim_data,mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_path(data=sim_data,mapping=aes(1000*y1,1000*y2),size=0.1)+
  geom_point(data = sim_data%>% filter(!duplicated(ID)),
             aes(x = y1* 1000, y = y2 * 1000), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggplotly(plot_sim)
```

### S

## Estimate from high frequency data without error: DistanceShore and tensor splines ti

```{r}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne2 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1,],type = "RACVM1",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne2$fit(method="BFGS")
```

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_rect_hf_ne2$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne2",npost=1000,level=0.95,
                          xmin=xmin,xmax=xmax)
res$plot_omega_theta_DistanceShore
```

Plot true and estimated omega as a function of theta for fixed values of DistanceShore, with confidence intervals.

```{r}
#plot spline estimates of the parameter omega close and far to the shore

predict_close=data.frame(theta=seq(-pi,pi,length.out=100),DistanceShore=rep(0.3,100))
predict_far=data.frame(theta=seq(-pi,pi,length.out=100),DistanceShore=rep(2,100))

sde_par_close <- crcvm_rect_hf_ne2$par(t = "all",new_data=predict_close)
sde_CI_close <- crcvm_rect_hf_ne2$CI_simultaneous(t = "all",new_data=predict_close)
sde_par_far <- crcvm_rect_hf_ne2$par(t = "all",new_data=predict_far)
sde_CI_far <- crcvm_rect_hf_ne2$CI_simultaneous(t = "all",new_data=predict_far)

# Data frame for point estimates
sde_par_df_close <- data.frame(theta=predict_close$theta,omega_estimates = sde_par_close[, "omega"])
sde_par_df_far<- data.frame(theta=predict_far$theta,omega_estimates = sde_par_far[, "omega"])

# Data frame for CIs
sde_ci_df_close<- data.frame(theta=predict_close$theta,
                          lowomega = sde_CI_close["omega", "low",],
                          upomega = sde_CI_close["omega", "upp",])
sde_ci_df_far<- data.frame(theta=predict_far$theta,
                          lowomega = sde_CI_far["omega", "low",],
                          upomega = sde_CI_far["omega", "upp",])

omega_df=data.frame(theta=predict_close$theta,omega_close=fomega(rep(0.3,100),predict_close$theta),omega_far=fomega(rep(2,100),predict_close$theta))

plot_omega_close=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_close) + 
  geom_ribbon(data=sde_ci_df_close,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  xlab("Theta")+
  geom_line(data=omega_df,aes(theta,omega_close),col="red")

plot_omega_far=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_far) + 
  geom_ribbon(data=sde_ci_df_far,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  xlab("Theta")+
  geom_line(data=omega_df,aes(theta,omega_far),col="red")

plot_omega_close
plot_omega_far
```

## Estimate from low frequency data without error : ExpShore and tensor splines ti

```{r}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_lf_ne1 <- SDE$new(formulas = formulas,data = data_rect_lf_ne[data_rect_lf_ne$ID==1,],type = "RACVM1",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),other_data=list(H=H0_lf))
crcvm_rect_lf_ne1$fit(method="BFGS")
```

```{r}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.1)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
crcvm_rect_lf_ne1$get_all_plots(baseline=NULL,model_name="crcvm_rect_lf_ne1",npost=1000,level=0.95,xmin=xmin,xmax=xmax,xlabel=xlabel,link=link)

```{r}
#plot spline estimates of the parameter omega close and far to the shore

predict_close=data.frame(theta=seq(-pi,pi,length.out=100),ExpShore=rep(1/0.3,100))
predict_far=data.frame(theta=seq(-pi,pi,length.out=100),ExpShore=rep(1/2,100))

sde_par_close <- crcvm_rect_lf_ne1$par(t = "all",new_data=predict_close)
sde_CI_close <- crcvm_rect_lf_ne1$CI_simultaneous(t = "all",new_data=predict_close)
sde_par_far <- crcvm_rect_lf_ne1$par(t = "all",new_data=predict_far)
sde_CI_far <- crcvm_rect_lf_ne1$CI_simultaneous(t = "all",new_data=predict_far)

# Data frame for point estimates
sde_par_df_close <- data.frame(theta=predict_close$theta,omega_estimates = sde_par_close[, "omega"])
sde_par_df_far<- data.frame(theta=predict_far$theta,omega_estimates = sde_par_far[, "omega"])

# Data frame for CIs
sde_ci_df_close<- data.frame(theta=predict_close$theta,
                          lowomega = sde_CI_close["omega", "low",],
                          upomega = sde_CI_close["omega", "upp",])
sde_ci_df_far<- data.frame(theta=predict_far$theta,
                          lowomega = sde_CI_far["omega", "low",],
                          upomega = sde_CI_far["omega", "upp",])

omega_df=data.frame(theta=predict_close$theta,omega_close=fomega(rep(0.3,100),predict_close$theta),omega_far=fomega(rep(2,100),predict_close$theta))

plot_omega_close=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_close) + 
  geom_ribbon(data=sde_ci_df_close,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  xlab("Theta")+
  geom_line(data=omega_df,aes(theta,omega_close),col="red")

plot_omega_far=ggplot() +geom_line(aes(theta, omega_estimates),data=sde_par_df_far) + 
  geom_ribbon(data=sde_ci_df_far,aes(x=theta,ymin=lowomega,ymax=upomega),alpha=0.2)+
  xlab("Theta")+
  geom_line(data=omega_df,aes(theta,omega_far),col="red")

plot_omega_close
plot_omega_far
```

## Estimate from low frequency data without error : DistanceShore and tensor splines ti

```{r}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_lf_ne2 <- SDE$new(formulas = formulas,data = data_rect_lf_ne[data_rect_lf_ne$ID==1,],type = "RACVM1",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_lf))
crcvm_rect_lf_ne2$fit(method="BFGS")
```

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.1)
xmax=list("theta"=pi,"DistanceShore"=2)
crcvm_rect_lf_ne2$get_all_plots(baseline=NULL,model_name="crcvm_rect_lf_ne2",npost=1000,level=0.95,
                          xmin=xmin,xmax=xmax)

## Estimate from high frequency data with error : ExpShore and tensor splines ti

```{r}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=5,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_e1 <- SDE$new(formulas = formulas,data = data_rect_hf_e[data_rect_hf_e$ID==1,],type = "RACVM1",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H1_hf))
crcvm_rect_hf_e1$fit(method="BFGS")
```

>>>>

---
title: "Simulation study in rectangular domain"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

@@

#used in function split_term to get covariates names from the formulas
library(stringr)

#to fit smooth function with tensor splines
library(mgcv)

#some auxiliary functions
source("CVM_functions.R")
```

# SIMULATION

## Create the rectangle

@@

#rectangular domain 
large_rectangle_coords <- matrix(c(0, 0, 0, 10, 10, 10, 10, 0, 0, 0), ncol = 2, byrow = TRUE)
small_rectangle_coords <- matrix(c(1,1, 1, 9, 9, 9, 9, 1, 1, 1), ncol = 2, byrow = TRUE)

@@

## Sample initial positions uniformly in the domain

@@

#number of tracks
n_samples=1

@@

## Fix time steps for observations

@@

hf=1/60 #1 minute time step

@@

@@

## Set observation error

```{r}
#no measurement error (we cannot put 0 because only the log of this quantity is used)
sigma_obs_0=0.005
H0_hf=array(rep(sigma_obs_0^2*diag(2),n_hf),dim=c(2,2,n_hf))
#GPS measurement error 
sigma_obs_1=0.03
H1_hf=array(rep(sigma_obs_1^2*diag(2),n_hf),dim=c(2,2,n_hf))

## Smooth parameters of the CRCVM

### Define smooth functions for tau and omega.

@@

ftau=function(cov_data,D0=0.3,delta0=0.2,tau0=1,tau1=2,tau2=2,lambda=4,kappa=0.1) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  coeff=exp(-kappa*(Dshore/D0)^2)
  tau=tau0*(1-coeff)+(delta0+(tau0-delta0)*exp(-theta^2/pi^2)+(tau1-delta0)*exp(-lambda*(theta-pi/2)^2)+(tau2-delta0)*exp(-lambda*(theta+pi/2)^2))*coeff
  return (tau)
}

#mixture of independent gaussian
ftau_gauss=function(cov_data,D0=0.5,tau0=1,tau1=1.5) {
  Dshore=cov_data$DistanceShore
  theta=cov_data$theta
  tau=tau0+(tau1-tau0)*exp(-((theta-pi/2)/(pi/10))^2-(Dshore/D0)^2)+(tau1-tau0)*exp(-((theta+pi/2)/(pi/10))^2-(Dshore/D0)^2)
  return (tau)
}

#constant tau
ftau_constant=function(cov_data,tau=1) {
  return (tau)
}

fomega=function(cov_data,D0=0.3,omega0=60*pi,lambda=1.5,kappa=0.5) {
    Dshore=cov_data$DistanceShore
    theta=cov_data$theta
    if (is.null(Dshore)){
      Dshore=1/cov_data$ExpShore
    }
    coeff=exp(-kappa*(Dshore/D0)^2)
    omega=omega0/2*(tanh(lambda*(theta+pi/2-atanh(-0.9)/lambda))+tanh(lambda*(theta-(pi/2-atanh(-0.9)/lambda))))*coeff
    return(omega)
}

@@

### Plot smooth functions.

Dshore=seq(from=0.3,to=2,length.out=30)
theta=seq(from=-pi,to=pi,length.out=30)
grid<- as.data.frame(expand.grid(Dshore,theta))
colnames(grid)=c("DistanceShore","theta")
true_par=grid
true_par$tau=ftau_constant(grid)
true_par$omega=fomega(grid)

@@

### Approximate smooth functions with tensor splines of DistanceShore

```{r}
sp_df=5
#number of points
n <- 500

#sample theta and DistanceShore points in the domain
theta <- runif(n,-pi,pi);
DistanceShore <- runif(n,0.2,2);
samples=data.frame(theta=theta,DistanceShore=DistanceShore)

#define grid of values
theta_v <- seq(-pi,pi,length=30)
Dshore_v<- seq(0.2,2,length=30)
pr <- data.frame(theta=rep(theta_v,30),DistanceShore=rep(Dshore_v,rep(30,30)))

# true values of the function over this grid
truth <- matrix(fomega(pr),30,30)

#points on the surface perturbed by gaussian noise
f <- fomega(samples)
y <- f+0*rnorm(n)

#plot true function
persp(theta_v,Dshore_v,truth);
title("truth")

#fit with bivariate splines te
m1 <- gam(y~te(theta,DistanceShore,k=sp_df))

#visualize
vis.gam(m1);title("tensor product")

df=data.frame("x"=theta,"z"=DistanceShore,"y"=y)
fig <- plot_ly(df,type="scatter3d", x = ~x, y = ~z, z = ~y,mode="markers",marker=list(size=4))
fig
```

### Approximate smooth functions with tensor splines of ExpShore

```{r}
ExpShore=1/DistanceShore

#fit with bivariate splines te
m2 <- gam(y~te(theta,ExpShore,k=sp_df))

#visualize
vis.gam(m2);title("tensor product")

```
### Get estimated spline coefficients

```{r}
sp_coeff_Dshore=m1$coefficients
sp_coeff_ExpShore=m2$coefficients
```

## Simulation with measurement errors and high frequency

### Generate sample
```{r, cache=TRUE,results='hide'}

@@

### percentage of samples that reached land

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_ne[data_rect_hf_ne$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### plot one sample

@@

plot_samples_rect_hf_ne=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_rect_hf_ne[data_rect_hf_ne$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_rect_hf_ne[data_rect_hf_ne$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
  xlab("x") + ylab("y")

@@

## Simulation with high measurement error and high frequency

### Generate samples

```{r, cache=TRUE,results='hide'}

@@

### percentage of samples that reached land

```{r}
count=0
for (id in unique(data_rect_hf_ne$ID)){
  sub_data=data_rect_hf_e[data_rect_hf_e$ID==id,]
  if (nrow(sub_data) < n_hf) {
    print(id)
    count=count+1
  }
}
cat(count/n_samples*100,"percent of the samples reached land")
```

### plot one sample

@@

plot_samples_rect_hf_e=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_path(data=data_rect_hf_e[data_rect_hf_e$ID==1,],mapping=aes(y1,y2),size=0.1)+
  geom_point(data = data_rect_hf_e[data_rect_hf_e$ID==1,]%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

@@

# ESTIMATION

## Compute distance to shore and empirical deviation angles

@@

data_rect_hf_ne=add_covs(data_rect_hf_ne)

data_rect_hf_e=add_covs(data_rect_hf_e)
```

## Estimate from high frequency data without low error

### Use ExpShore covariate and tensor splines te

#### Fit sde

```{r,cache=TRUE}

#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,ExpShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne1 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne1$fit(method="BFGS")
```

#### Plot estimated smooth parameter omega in 3D

```{r}
xmin=list("theta"=-pi,"ExpShore"=1/2)
xmax=list("theta"=pi,"ExpShore"=1/0.3)
xlabel=list("ExpShore"="Distance to shore")
link=list("ExpShore"=(\(x) 1/x))
true_smooths=list("omega"=fomega)
res=crcvm_rect_hf_ne1$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne1",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_ExpShore
```

#### Violin plot of the estimated spline coefficients

@@

true_coeff=data.frame(value=sp_coeff_ExpShore,name=paste("omega.",names(sp_coeff_ExpShore),sep=""))
rownames(true_coeff)=NULL

p=crcvm_rect_hf_ne1$plot_violin_coeff(model_name="crcvm_rect_hf_ne1",substring="omega",true_coeff=true_coeff,save=TRUE)

ggplotly(p)
```

#### Simulate from fitted model

```{r,cache=TRUE}

@@

atw=list("ExpShore"=fExpShore,"AngleNormal"=fangle)
sim_data=crcvm_rect_hf_ne1$simulate(z0=x0[1,],data=crcvm_rect_hf_ne1$data(),atw=atw,land=domain)

@@

#### Plot simulation from fitted model

@@

plot_sim=ggplot()+geom_sf(data=domain$geometry,fill=NA)+coord_sf(datum=st_crs(32626))+
  geom_point(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_path(data=sim_data,mapping=aes(y1,y2),size=0.1)+
  geom_point(data = sim_data%>% filter(!duplicated(ID)),
             aes(x = y1, y = y2 ), shape = 3, size = 4, col = "red")+
    xlab("x") + ylab("y")

ggsave("sim_crcvm_rect_hf_ne1.png",plot=plot_sim,width=10,height=5,path="crcvm_rect_hf_ne1")
ggplotly(plot_sim)
```

###  UseDistanceShore covariate and tensor splines ti

#### Fit sde

```{r,cache=TRUE}
#estimation with tensor splines
formulas <- list(mu1=~1,mu2=~1,tau=~1,
                 nu=~1,omega=~te(theta,DistanceShore,k=sp_df,bs="cs"))
par0 <- c(0,0,1,4,0)

crcvm_rect_hf_ne2 <- SDE$new(formulas = formulas,data = data_rect_hf_ne[data_rect_hf_ne$ID==1,],type = "RACVM",
                      response = c("y1","y2"),par0 = par0,fixpar=c("mu1","mu2","nu","tau"),
                      other_data=list("H"=H0_hf))
crcvm_rect_hf_ne2$fit(method="BFGS")
```

#### Plots estimated smooth in 3D

```{r}
xmin=list("theta"=-pi,"DistanceShore"=0.3)
xmax=list("theta"=pi,"DistanceShore"=2)
res=crcvm_rect_hf_ne2$get_all_plots(baseline=NULL,model_name="crcvm_rect_hf_ne2",xmin=xmin,xmax=xmax,
                                    xlabel=xlabel,link=link,show_CI="pointwise",true_smooths=true_smooths,save=TRUE)
res$fe_omega_theta_DistanceShore
```

#### Violin plot of the estimated spline coefficients
